\documentclass[letterpaper, 10pt]{article}

\usepackage[left=.5in, top=.25in, total={7.5in, 10.5in}, includeheadfoot]{geometry}

\usepackage{booktabs}

\usepackage{tabularx}

\usepackage{longtable}

\usepackage{array}

\usepackage{hyperref}     % for clickable links

\renewcommand{\arraystretch}{1.25}



% Custom command

%\newcommand{\headings}[1]{\section*{#1} \hrule \vspace{10pt}}
\newcommand{\headings}[1]{\section*{\textsf{#1}} \hrule \vspace{10pt}}
\newcommand{\headingstwo}[1]{\section*{#1} \vspace{-20pt}}

\title{CV}
\author{bhowes }
\date{April 2026}

\begin{document}
\begin{center}
    {\LARGE \textbf{Bolton Howes}} \\[6pt]
    \textit{Assistant Professor \& Director of the Black Hills Field Station} \\[3pt]
    Wheaton College \\[8pt]
    
    % Contact line with separators
    \href{mailto:bolton.howes@wheaton.edu}{bolton.howes@wheaton.edu} $|$ 
    \href{https://boltonhowes22.github.io}{boltonhowes22.github.io}
\end{center}


\headings{Professional Experience}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
    2025-- & \textbf{Wheaton College} \newline Assistant Professor \& Director of The Field Station in the Black Hills\\	
    2024--2025 & \textbf{University of Washington \& Western Washington University} \newline NSF Postdoctoral Fellow %\newline Leveraging remote sensing, data science, and computational methods to understand how rivers respond to climate change
       \newline Advisors: Akshay Mehra and Brady Foreman\\	
  2023--2024 & \textbf{University of Miami (Rosenstiel School of Marine, Atmospheric, and Earth Science)} \newline Postdoctoral Researcher, Marine Geology
  %\newline Remote sensing of sedimentology to determine the natural hazard risk of the Red Sea
	\newline Advisor: Sam Purkis \\
\end{tabular}
\end{flushleft}

\headings{Education}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
		2017--2023 & \textbf{Princeton University}, Princeton, NJ \newline Ph.D., Geosciences \newline Dissertation: Developing physical paleoenvironmental proxies: from ooids to extinction
		\newline Advisor: Adam Maloof \\
		2015--2017 & \textbf{University of Georgia}, Athens, GA \newline M.Sc., Geology %\newline Thesis: The application of sequence stratigraphic principles to constrain the onset of thrust load-induced subsidence: the Jurassic Twin Creek Limestone of Southwest Wyoming
		\newline Advisor: Steven Holland\\
		2011--2015 & \textbf{Macalester College}, St. Paul, MN \newline B.A., Geology %\newline Thesis: Characterization of a regionally significant terrestrial bounding surface in the Upper Cretaceous Two Medicine Formation, northwestern Montana \newline Advisor: Raymond Rogers\\
	\end{tabular}
\end{flushleft}


\headings{Awards and Honors}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
        2024 & UW eScience Postdoctoral Fellowship [\$4,000]\\
        2023 & NSF EAR Postdoctoral Fellowship [\$180,000]\\
        2017--2022 & Department of Geosciences Graduate Research Fund [3x\$5,000]\\
		2017 & Master's Student of the Year, University of Georgia\\
		2015--2017 & Graduate Student Assistantship. University of Georgia (fully-funded master's program)\\
		2016 & Graduate Research Grant, Geological Society of America\\
		2016 & Miriam Watts-Wheeler Award, University of Georgia [\$1,600]\\
		2015 & Henry Lepp Award for Dedication to Scientific Research, Macalester College\\
	\end{tabular}
\end{flushleft}



\headings{Publications}
%\noindent \textit{Submitted \& In Review}
% Global list spacing (applied via enumitem or manually in each list)


\noindent \textit{In review}
\begin{enumerate}
    \item[14.] Morgan Chakraborty, Hannah Shernisky, Clement Pllier, Amanda Oehlert, Arash Sharifi, Ali Pourmand, \textbf{Bolton Howes}, Benjamin Minch, Mohammad Moniruzzaman, Larissa Fruhe, Mattie Rodrigue, Sam Purkis. Brine Pool Microbes Enrich Critical Metals in Salt Giant Basins. AGU Advances
    \item[13.] Anthony Semerraro, Brian Cardenas, \textbf{Bolton Howes}, Brady Foreman. Reconstructing river planform morphology in alluvial stratigraphy using paleocurrents.

\end{enumerate}

\clearpage

\noindent \textit{In prep.}

\begin{enumerate}
	\item[12.] Filip Novak, \textbf{Bolton Howes}, Henry Yuan, Brady Foreman, Akshay Mehra. From outcrop to point cloud: A three-dimensional technique for the remote analysis of terrestrial and extraterrestrial sedimentary structures
    
	\item[11.]  \textbf{Bolton Howes}, Morgan Chakraborty, Sam Purkis. The afterlife of lowstand sediments.
    
    \item[10.] \textbf{Bolton Howes}, Delaney Todd, Katherine Sullivan, Akshay Mehra, Brady Foreman. Seasonality and discharge variability cause increased river mobility and floodplain erosion during the Paleocene-Eocene global warming event
\end{enumerate}


\noindent \textit{Published}
\begin{enumerate}
    \item[9.] Kevin I. Velez-Rosado, Olga Zalles-Grebetskaya, Jeffrey A. Wilson Mantilla, Blair Schoene, Adam Maloof, \textbf{Bolton Howes}. New material of \textit{Dolichochampsa minima} (Archosauria: Crocodylia) from the Cretaceous--Paleogene El Molino Formation of Bolivia sheds light on the early evolution of Gavialinae. 2025. \textit{Journal of Systematic Paleontology}.

    \item[8.] Morgan Chakraborty, Francesca Benzoni, François Tissot, Ali Pourmand, Marco Taviani, \textbf{Bolton Howes}, Peter Swart, Mattie Rodrigue, Sam Purkis. Deep-water corals indicate the Red Sea survived the last glacial lowstand. 2025. \textit{PNAS}.
    
    \item[7.] Sam Purkis, Steve Ward, \textbf{Bolton Howes}, Morgan Chakraborty, Jake Longenecker. A 1600-year record of extreme rainfall in northern Arabia. 2025. \textit{Science Advances}.
  
    \item[6.] \textbf{Bolton Howes}, Akshay Mehra, Emily Geyman, Julia Wilcots, Ryan Manzuk, Curtis Deutsch, Adam Maloof. The where, when, and how of ooid formation: what ooids tell us about ancient seawater chemistry. 2025. \textit{Earth and Planetary Science Letters}.

    \item[5.] Gaelle Duchâtellier, Amanda M. Oehlert, Hannah Shernisky, Clement G.L. Pollier, Peter K. Swart, \textbf{Bolton Howes}, Sam J. Purkis. Sedimentary porewaters record regional tectonic and climate events that perturbed a deep-sea brine pool in the Gulf of Aqaba, Red Sea. 2024. \textit{ Science of the Total Environment}.

    \item[4.] Akshay Mehra, \textbf{Bolton Howes}, Ryan Manzuk, Alec Spatzier, Bradley Samuels, Adam Maloof. A novel technique for producing three-dimensional data using serial sectioning and semi-automatic image classification. 2022. \textit{Microscopy and Microanalysis}.

    \item[3.] Emily Geyman, Ziman Wu, Matthew Nadeau, Stacey Edmonsond, Andrew Turner, Sam Purkis, \textbf{Bolton Howes}, Blake Dyer, Anne-Sophie Ahm, Nan Yao, Curtis Deutsch, and Adam Maloof. The origin of carbonate mud and implications for global climate. 2022. \textit{PNAS}.

    \item[2.] \textbf{Bolton Howes}, Akshay Mehra, and Adam Maloof. Three-Dimensional Morphometry of Ooids in Oolites: a new tool for more accurate and precise paleoenvironmental interpretation. 2021. \textit{Journal of Geophysical Research: Earth Surface}.

    \item[1.] Paul Hoffman, Galen Halverson, Daniel Schrag, John Higgins, ... \textbf{Bolton Howes}. Snowballs in Africa: sectioning a long-lived Neoproterozoic carbonate platform and its bathyal foreslope (NW Namibia). 2021. \textit{Earth-Science Reviews}.
\end{enumerate}



\headings{Conference Abstracts \textnormal{\footnotesize(* denotes student advisee)}} 
\begin{flushleft}
	\begin{longtable}  
    {p{1.75cm}|p{16cm}}
        2025 & Filip Novak\textsuperscript{*}, \textbf{Bolton Howes}, Brady Foreman, Akshay Mehra. From Outcrop to Point Cloud: Using Three-Dimensional Techniques to Study Fluvial Dynamics During the Paleocene-Eocene in Wyoming. \textit{GSA} (2025).\\

        2025 & \textbf{Bolton Howes}, Anthony Semerraro, Akshay Mehra, Brady Foreman. Reconstructing river planform morphology. \textit{Geobiology} (2025).\\
         
        2022 & \textbf{Bolton Howes}, Ryan Manzuk, Akshay Mehra, Adam Maloof. Tales of transport and chemistry from three-dimensional growth histories of ooids. \textit{AGU Fall Meeting} (Chicago, Illinois).\\

        2022 & Ryan Manzuk, \textbf{Bolton Howes}, Emily Geyman, Indu Panigrahi, Dave Singh, Adam Maloof. Multispectral petrography image analysis as a method for paleoenvironmental and paleoecological reconstruction; applications to Earth's first reefs. AGU Fall Meeting (Chicago, Illinois).\\

        2022 & Cedric Hagen, John Higgins, Adam Maloof, Stacey Edmonsond, Ryan Manzuk, \textbf{Bolton Howes}, Pingping Zhao. Machine learning and computational image analysis leveraged to probe relationships between carbonate isotopic composition, diagenetic alteration, and sedimentary facies isotopic variability. AGU Fall Meeting (Chicago, Illinois).\\
        
        2022 & Adam Maloof, Ryan Manzuk, Emily Geyman, Akshay Mehra, Jaap Kaandorp, Mark Webster, Stacey Edmonsond, \textbf{Bolton Howes}, and Cedric Hagen. From modern analogs to three dimensions: Lessons learned for interpreting the stratigraphic record of the Proterozoic--Phanerozoic transition. GSA Annual Meeting (Denver, Colorado).\\

        2018 & \textbf{Bolton Howes}, Akshay Mehra, and Adam Maloof. Three-dimensional reconstructions of Holocene and Neoproterozoic ooltes to measure porosity, permeability, and volume-shape evolution of ooids. GSA Annual Meeting (Indianapolis, Indiana).\\
        
        2014 & \textbf{Bolton Howes} and Raymond Rogers. Revisiting a regionally significant terrestrial bounding surface in the Upper Cretaceous (Campanian) Two Medicine Formation, northwestern Montana. GSA Annual Meeting (Vancouver, British Columbia).
    \end{longtable}
\end{flushleft}

\headings{Invited Talks}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{13cm}}
        2025 & Purdue University\\
        2025 & University of Washington: PaleoLunch\\
        2025 & Macalester College, Sed/Strat Seminar\\
        2025 & Western Washington University, Department Seminar\\
        2024 & Wheaton College, Department Seminar\\
	2020 & Macalester College, Graduate School Open House\\
	2018 & Macalester College, Sed/Strat Seminar\\
	\end{tabular}
\end{flushleft}

\headings{Fieldwork Experience (82 Weeks)}
\begin{flushleft}
	\begin{longtable}{p{1.75cm}|p{2.75cm}p{1.5cm}p{11cm}}
		2026 & Bolivia & 2 weeks & K-Pg Paleontology\\	    
		2024 & Indonesia & 2 weeks & Mapping seafloor ecology and natural hazards\\	    
        \fontsize{9}{13}\selectfont 2024 \footnotesize{\&} \fontsize{9}{13}\selectfont 2025 & NW Wyoming & 5 weeks & The fluvial response to the PETM\\
        2023 & Red Sea & 3 weeks & Geomorphology and stratigraphy to evaluate natural hazard risk\\
        \fontsize{9}{13}\selectfont 2019 \footnotesize{\&} \fontsize{9}{13}\selectfont 2022 & Bolivia & 12 weeks & The K--Pg Boundary in lacustrine and fluvial settings\\
	    2021 & southern Utah & 2 weeks & Cyclicity in the Late Paleozoic Ice Age \\
	    2019 & Calabria, Italy & 1 week & Measuring climate change in olive orchards \\
	    2019 & Yukon, Canada & 5 weeks & Archeocyathid reefs and the Cambrian Explosion\\
	    2018 & Spain and Italy & 6 weeks & The K--Pg Boundary and slope sediments \\
	    2018 & southern Namibia & 4 weeks & Neoproterozoic and Paleozoic glacial deposits \\
		2018 & Western Australia & 1 week & Modern and recent carbonate platforms \\
		2018 & Nevada/Arizona & 3 weeks & Stratigraphy and Geochemistry of the late Paleozoic ice age \\
		2017 & northern Namibia & 6 weeks & Field assistant to Paul Hoffman. Mapping Snowball Earth\\
		2016 & western Wyoming & 6 weeks & Sequence strat. of the Twin Creek Limestone\\
		2014 & NW Montana & 2 weeks & Sequence strat. and strat. paleobiology of the Two Medicine Formation\\
		2014 & Patagonia, Chile & 12 weeks & Restoration ecology and conservation biology in the Aysén Region\\
		\fontsize{9}{13}\selectfont 2013 \footnotesize{\&} \fontsize{9}{13}\selectfont 2014 & NW Wyoming & 5 weeks & Terrestrial stratigraphy and paleobiology of the Cloverly Formation \\
		2014 & Tall Dhiban, \newline Jordan & 6 weeks & Archeological excavation  Early Bronze Age through the 19th century \\
	\end{longtable}
\end{flushleft}



\headings{Undergraduates Advised}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{3.5cm}p{12cm}}
			2026 & Carson Frear & Bolivian K--Pg\\
            2024--2025 & Filip Novak & Photogrammetry and fluvial sedimentology \\ 
            2021 & Anubhav Agarwal & Transfer learning for geological image segmentation  \\
            2021 & Hugh Shields & Transfer learning for geological image segmentation\\
            2021 & Anna Krokhine & Numerical models of ooid growth\\
            2020 & Emilio Cano-Renteria & Geological image segmentation and ooid growth\\
            2019 & Sarah Brown & K--Pg boundary stratigraphy of Bolivia\\
		2018 & Celia Aranda-Reina & Climate variability during the K--Pg in Spain\\
	\end{tabular}
\end{flushleft}


\headings{Undergraduates Involved in Fieldwork}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{3.5cm}p{1.5cm}p{8cm}}
			
			2026 & Carson Frear & 2 weeks & Bolivia\\
            2025 & Filip Novak & 2 weeks & Wyoming\\
            2019 & 10 Undergraduates & 1 week & Italy\\
            2019 & Sarah Brown & 5 weeks & Bolivia\\
		2019 & Galen Cadley  & 5 weeks & Yukon, Canada\\
            2019 & 8 Undergraduates & 1 week & Utah\\
		2018 & Celia Aranda-Reina & 4 weeks & Spain\\
		2016 & Tony Moraes & 4 weeks & western Wyoming\\
	\end{tabular}
\end{flushleft}

\clearpage

\headings{Teaching}

\noindent \textit{Instructor of Record}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
        % ADD YOUR WHEATON COURSES HERE. Examples:
        Spr. 2026 & \textbf{GEOL 321:} Earth History and Stratigraphy, Wheaton College\\
	\end{tabular}
\end{flushleft}

\noindent \textit{Teaching Assistant}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
	    Spr. 2021 & \textbf{ENV 354-GEO 368} Climate and Weather: Order in the Chaos, Princeton University\\
	    Fall 2019 $^{\textbf{*}}$  & \textbf{FRS 161:} Earth: Crops, Culture, and Climate (in Italy), Princeton University\\
		Fall 2018 $^{\textbf{*}}$ & \textbf{GEOL 201:} Measuring Climate Change: Data Analysis and Science Writing, Princeton University\\
		2016--2017 & Geology Tutor, University of Georgia Student-Athlete Academic Center\\
		Spr. 2015 & \textbf{GEOL 165}: History and Evolution of the Earth, Macalester College\\
		Fall 2014 & \textbf{GEOL 300}: Paleobiology, Macalester College\\
	\end{tabular}
\end{flushleft}

\noindent\textit{$^{\textbf{*}}$ Indicates courses that included a fieldwork component consisting of a week-long trip (Utah and Italy) to make observations for research projects. My teaching duties included preparing supplies and logistics, and helping students with both data collection---especially with drones, differential GPS, and surveying equipment---and data analysis. The data collected during fieldwork was part of a semester-long project. }



\headings{Leadership and Service}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{13cm}}
 	    2023 & SIFP Fellow (mentoring first-gen. students), Princeton University\\
	    2018--2021 & Residential Graduate Student, Princeton University\\
		2016--2017 & Student Volunteer, Geological Society of America\\
		2015 & Leadership Committee of Macalester College Fellowship of Christian Athletes\\ 
		2012--2013 & Resident Assistant, Macalester College\\
		2011--2014 & Captain of Macalester College Football Team\\

	\end{tabular}
\end{flushleft}


\headings{Professional Activities}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{13cm}}
 	    Convener & 2019 GSA Annual Meeting technical session: Hello (ancient) world!: Exploring the Neoproterozoic to Cambrian interval by quantitatively probing the rock record\\
	    Reviewer & NSF EAR Geoinformatics, Geophysical Research Letters\\
	\end{tabular}
\end{flushleft}


\headings{Public Engagement}
\begin{flushleft}
	\begin{tabular}{p{1.75cm}|p{15cm}}
        2025 & \textit{Tiny Living Beings} podcast (American Society for Microbiology): ``4.5 Billion Years of Earth History in One Hour''\\
        % ADD MORE HERE: outreach talks, science communication, K-12 visits, etc.
	\end{tabular}
\end{flushleft}




%\headings{Continued Education}
%\begin{flushleft}
%    \begin{tabular}{p{1.75cm}|p{15cm}}
%		2018 & Agouron Advanced Astrobiology and Geobiology Field School in Namibia, Agouron Institute\\
%		2016 & Reservoir Engineering, ExxonMobil\\
%		2016 & Geochemistry of the Cretaceous Western Interior Seaway, Geological Society of America\\
%		2015 & Sequence Stratigraphy for Graduate Students, ExxonMobil\\
%	\end{tabular}
%\end{flushleft}

%\headings{References}
%\begin{center}
%	\begin{tabularx}{\textwidth}{>{\itshape\arraybackslash}l X}
%		Adam Maloof & maloof@princeton.edu\\
%		Ray Rogers & rogers@macalester.edu\\
%            Ian Graham (Pastor) & ian@ecclesianj.com\\
%	\end{tabularx}
%\end{center}


\end{document}
